Generality in food webs scales with species richness, not latitude

### 2014-07-09 manuscript/latitudepaper.tex
Clean enough to pass along.

%Food webs are a thing. (Mention gen, vul, LS.) %Food webs have strong scaling structure.
Food webs (networks of feeding links between species) have been used for decades to summarize the structure of 
ecological communities \citep{Williams2000,earlierwork}.
One of the earliest goals of constructing food webs was to measure the relationship between the species richness of
a community and its stability \citep{Dunne2006}, a goal which is still pursued today \citep{robustness studies}.
Some authors have recognized the potential for environmental effects to influence both stability \citep{Shurin2007} and 
other food web properties such as food chain length \citep{Townsend1998,Thompson2003,2004c,Young2013,Digel2014}.
Latitude is a particularly intruiging contender for an environmental influence on food web structure \citep{Shurin2007},
as latitudinal gradients in temperature \citep{}, primary productivity \citep{}, body size \citep{}, and 
species richness \citep{Schemske2009a} are among the most robust patterns in ecology[[back up from original draft]].

Latitude's strong influence on species richness, however, presents a problem for the analysis of food web structure.
Although it was initially thought that food-web structure was small-world and scale invariant \citep{}, further 
analyses with higher-quality data have revealed that most food-web structural properties scale strongly with species 
richness, connectance (the proportion of possible links that are realized), or both \citep{Riede2010} 
(see \citet{Dunne2006} for a review of the history of scaling laws in food web theory). This means that relationships
between latitude and food-web structure must be considered in light of the latitudinal gradient in species richness.
This is especially true for food-web properties that both scale with species richness and are related to candidate
explanations for the latitudinal diversity gradient, such as measures of specialization.

@@

S_{i} = \alpha_{0} + \alpha_{1}\Delta_{i}+\epsilon_{i} \,,

@@

The tendency of food-web structure to exhibit scaling relationships with both
species richness and connectance has been  well-established \citep{,,,,}. As
species richness in particular is also known to vary systematically over
latitude  \citep{,,,,}, intuitively one might suspect that any relationship
between food-web properties such as generality might be due to the latitudinal
gradient in species richness. Here, we found that while there were latitudinal
gradients in species richness and generality, latitude was not related to
either the scaling relationship between generality and species richness or the
residuals of this scaling relationship.

As the latitudinal gradient in species richness in this data set ran in the
opposite direction to the general trend for angiosperms \citep{}, mammals
\citep{}, birds \citep{}, marine organisms? \citep{}, we conclude that it is
likely an artifact of the data set. Academic institutions have the longest
history in temperate latitudes, and it may be that communities at these
latitudes are better understood or more completely sampled. We do not believe
that our observation of higher species richness at higher latitudes represents
a true biological trend. As such, it does not seem likely that the increase in
generality with latitude that we observed is due to an underlying tendency for
species to pursue a wider variety of prey at higher latitudes. Rather, the
trend in generality can be more parsimoniously explained by the trend in
species richness.

>>>>

%Food webs are a thing. (Mention gen, vul, LS.) %Food webs have strong scaling structure.
Food webs (networks of feeding links between species) have been used for decades to summarize the structure of 
ecological communities \citep{Williams2000,earlierwork} and to understand how that structure relates to environmental variables 
such as land use \citep{Townsend1998,Townsend2004c,Digel2014}, primary productivity \citep{}, temperature?, latitude?. 
Latitude is of particular interest as an environmental driver of food-web structure because it entails gradients in 
temperature \citep{}, primary productivity \citep{}, and seasonality \citep{}. Importantly, there is also a strong
latitudinal gradient in species richness for most taxa \citep{Schemske2009a}.

This effect of latitude on species richness, however, presents a problem for the analysis of food web structure.
Although it was initially thought that food-web structure was small-world and scale invariant \citep{}, further 
analyses with higher-quality data have revealed that most food-web structural properties scale strongly with species 
richness, connectance (the proportion of possible links that are realized), or both \citep{Riede2010} 
(see \citet{Dunne2006} for a review of the history of scaling laws in food web theory). This means that relationships
between latitude and food-web structure must be considered in light of the latitudinal gradient in species richness.
This is especially true for food-web properties that both scale with species richness and are related to candidate
explanations for the latitudinal diversity gradient, such as measures of specialization.

@@

$S_{i} = \alpha_{0} + \alpha_{1}\Delta_{i}+\epsilon_{i}$ \,,

@@

The tendency of food-web structure to exhibit scaling relationships with both
species richness and connectance has been  well-established \citep{}. As
species richness in particular is also known to vary systematically over
latitude \citep{}, intuitively one might suspect that any relationship
between food-web properties such as generality might be due to the latitudinal
gradient in species richness. Here, we found that while there were latitudinal
gradients in species richness and generality, latitude was not related to
either the scaling relationship between generality and species richness or the
residuals of this scaling relationship.

As the latitudinal gradient in species richness in this data set ran in the
opposite direction to the general trend for \citep{}, we conclude that it is
likely an artifact of the data set. Academic institutions have the longest
history in temperate latitudes, and it may be that communities at these
latitudes are better understood or more completely sampled. As such, it does not seem likely that the increase in
generality with latitude that we observed is due to an underlying tendency for
species to pursue a wider variety of prey at higher latitudes. Rather, the
trend in generality can be more parsimoniously explained by the trend in
species richness.
